Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-07-16 MSU-thesis-estee.tex
backup
Bill edits added Experiment section
Bill's edits and mine to results
Fixed Appendix
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

% This is a template for use with the MSU Thesis class
% Vesion 2.8 2017/12/13
%
% Class options: 
%[PhD]	Doctor of Philosophy (default) 
%[DEd]	Doctor of Education
%[DMA]	Doctor of Musical Arts
%[MA]	Master of Arts
%[MS]	Master of Science
%[MAT]	Master of Arts for Teachers
%[MBA]	Master of Business Administration
%[MFA]	Master of Fine Arts
%[MIPS]	Master of International Planning Studies
%[MHRL]	Master of Human Resources and Labor Relations 
%[MMus]	Master of Music
%[MPH]	Master of Public Health
%[MPP]	Master of Public Policy
%[MSW]	Master of Social Work
%[MURP]	Master in Urban and Regional Planning 
%%
% Default is PhD
%
%
% This template has everything in the right order.
% Just add real content and you're done!
%
\documentclass[]{msu-thesis}
%
% for a prettier, but possibly non-compliant table of contents use the [mixedtoc] option
% for a plain table of contents use the [plaintoc] option
% for a horrendous looking, but possibly required table of contents, use the [boldtoc] option
%
% If you have large tables/figures that need to be in landscape mode, add the [lscape] option

@@

% This is standard fontenc/inputenc for pdflatex
% If you use LuaLaTeX or XeLaTeX you should replace this with the fontspec package
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{overpic}
\usepackage{contour}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{mathtools}
\usepackage{comment}\renewcommand{\arraystretch}{1.5}
\usepackage[version=3]{mhchem}

@@

%
% Load any extra packages here
%
% You must specify the title of your thesis, your name, the field of study (not department), and the year
\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
\author{Justin Estee}
\fieldofstudy{Physics} % This should be in sentence case
\date{2020}

% If you want a dedication page, specify the text of the dedication here and uncomment the next command.
%
\dedication{This thesis is dedicated to someone.}
%
%User defined commands
\newcommand{\spirit}{S$\pi$RIT }
\newcommand{\numerr}[3]{ \num[parse-numbers=false]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
%\newcommand{\numerr}[3]{ \num[input-close-uncertainty]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }

@@

\DeclareSIUnit\torr{Torr}
\DeclareSIUnit\eVperc{\eV\per\clight}
\DeclareSIUnit\MeVA{\mega\eV A}
\DeclareSIUnit\clight{\text{\ensuremath{c}}}
\DeclareSIUnit\micron{\micro\metre}
\DeclareSIUnit\mrad{\milli\rad}
\DeclareSIUnit\gauss{G}

\includeonly{
chapters/ch_intro,
%ch_theory,
%chapters/ch_experiment,
%chapters/ch_data_analysis,
%chapters/ch_data_analysis_2,
chapters/ch_results
}

@@

% All the stuff before your actual chapters is called the front matter
\frontmatter
% First make the title page
\maketitlepage
% Next make the abstract
\begin{abstract}
% Your abstract goes here.  Master's 1 page max. PhD 2 page max.
\end{abstract}

@@

% If you have a dedication page, uncomment the next command to print the dedication page
%
%\makededicationpage
%
\clearpage
% Your Acknowledgements are formatted like a chapter, but with no number
\chapter*{Acknowledgements}
\DoubleSpacing % Acknowledgements should be double spaced
Your acknowledgements here.
%
\clearpage
% We need to turn single spacing back on for the contents/figures/tables lists
\SingleSpacing
\tableofcontents* % table of contents will not be listed in the TOC
\clearpage
\listoftables % comment this out if you have no tables
\clearpage
\listoffigures % comment this out if you have no figures
%
% If you have a list of abbreviations/symbols it would go here preceded by a \clearpage
% See the class documentation and the Memoir manual for how to create other lists 
%
\mainmatter
%
% The next line removes the dots in chapter headings in the TOC
% May violate thesis office rules
%\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}

\include{chapters/ch_intro}
%\include{chapters/ch_theory}
%\include{chapters/ch_experiment}
%\include{chapters/ch_data_analysis}
%\include{chapters/ch_data_analysis_2}
\include{chapters/ch_results}

@@

\appendix
\input{chapters/ch_appendix}

>>>>

% This is a template for use with the MSU Thesis class
% Vesion 2.8 2017/12/13
%
% Class options: 
%[PhD]	Doctor of Philosophy (default) 
%[DEd]	Doctor of Education
%[DMA]	Doctor of Musical Arts
%[MA]	Master of Arts
%[MS]	Master of Science
%[MAT]	Master of Arts for Teachers
%[MBA]	Master of Business Administration
%[MFA]	Master of Fine Arts
%[MIPS]	Master of International Planning Studies
%[MHRL]	Master of Human Resources and Labor Relations 
%[MMus]	Master of Music
%[MPH]	Master of Public Health
%[MPP]	Master of Public Policy
%[MSW]	Master of Social Work
%[MURP]	Master in Urban and Regional Planning 
%%
% Default is PhD
%
%
% This template has everything in the right order.
% Just add real content and you're done!
%
\documentclass[]{msu-thesis}

%
% for a prettier, but possibly non-compliant table of contents use the [mixedtoc] option
% for a plain table of contents use the [plaintoc] option
% for a horrendous looking, but possibly required table of contents, use the [boldtoc] option
%
% If you have large tables/figures that need to be in landscape mode, add the [lscape] option

@@

% This is standard fontenc/inputenc for pdflatex
% If you use LuaLaTeX or XeLaTeX you should replace this with the fontspec package
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{overpic}
\usepackage{contour}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{mathtools}
\usepackage{appendix}
\usepackage{comment}\renewcommand{\arraystretch}{1.5}
\usepackage[version=3]{mhchem}

@@

%
% Load any extra packages here
%
% You must specify the title of your thesis, your name, the field of study (not department), and the year
\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
\author{Justin Brian Estee}
\fieldofstudy{Physics} % This should be in sentence case
\date{2020}

% If you want a dedication page, specify the text of the dedication here and uncomment the next command.
%
\dedication{}
%
%User defined commands
\newcommand{\spirit}{S$\pi$RIT }
\newcommand{\numerr}[3]{ \num[parse-numbers=false]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
%\newcommand{\numerr}[3]{ \num[input-close-uncertainty]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }

@@

\DeclareSIUnit\torr{Torr}
\DeclareSIUnit\eVperc{\eV\per\clight}
\DeclareSIUnit\MeVA{A \mega\eV }
\DeclareSIUnit\clight{\text{\ensuremath{c}}}
\DeclareSIUnit\micron{\micro\metre}
\DeclareSIUnit\mrad{\milli\rad}
\DeclareSIUnit\gauss{G}

\includeonly{
chapters/ch_intro,
%ch_theory,
chapters/ch_experiment,
chapters/ch_data_analysis,
chapters/ch_data_analysis_2,
chapters/ch_results,
chapters/ch_summary,
chapters/ch_appendix
}

@@

% All the stuff before your actual chapters is called the front matter
\frontmatter
% First make the title page
\maketitlepage
% Next make the abstract
\begin{abstract}
% Your abstract goes here.  Master's 1 page max. PhD 2 page max.

Recent astronomical measurements of neutron star mergers have reinvigorated questions about the nature of dense matter in the universe. The densities reached in these stars range from  normal nuclear densities $\rho_o$ to upwards of $9\rho_o$, where $\rho_o = \SI{1.7e14}{\gram\per\centi\metre\squared}$, extremely dense matter, which is primarily composed of neutrons. The symmetry energy describes the difference in binding energy between pure neutron matter and symmetric matter with equal number of neutrons and protons. The density dependence of the symmetry energy is related to the pressure, which supports the star preventing it from collapsing under its own gravitation force. For years, theory and experiment  have combined to progressively constrain the equation of state of symmetric matter, and it was in the last decade that the density dependence of pure neutron matter (the symmetry energy), had been constrained at normal nuclear densities and lower, $\rho < \rho_o$. Still large uncertainties still remain at densities around 2$\rho_o$ which are important to the understanding the properties of neutron stars.

Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities, and to probe different proton-neutron asymmetries. Charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonance, in the early, dense parts of heavy pion collisions. Yet, existing experimental data sensitive to the symmetry energy was lacking, prompting the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pion spectra ($\pi^\pm$) with a high efficiency produced in neutron-rich heavy-ion collisions. This dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems. Also we will compare the results with the most recent transport  simulations. We will highlight some of the successes and hope to motivate a discussion in the theoretical community on how to better reproduce the pion phenomena in neutron-rich heavy-ion collisions, which is important for not only constraining the density dependence of the symmetry energy, but also for understanding the $\Delta$ and pion roles in neutron stars.

\end{abstract}

@@

% If you have a dedication page, uncomment the next command to print the dedication page
%
%\makededicationpage
%
\clearpage
% Your Acknowledgements are formatted like a chapter, but with no number
\chapter*{Acknowledgements}
\DoubleSpacing % Acknowledgements should be double spaced
Your acknowledgements here.
%
\clearpage
% We need to turn single spacing back on for the contents/figures/tables lists
\SingleSpacing
\tableofcontents* % table of contents will not be listed in the TOC
\clearpage
\listoftables % comment this out if you have no tables
\clearpage
\listoffigures % comment this out if you have no figures
%
% If you have a list of abbreviations/symbols it would go here preceded by a \clearpage
% See the class documentation and the Memoir manual for how to create other lists 
%
\mainmatter
%
% The next line removes the dots in chapter headings in the TOC
% May violate thesis office rules
%\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}
\chapterstyle{default}

\include{chapters/ch_intro}
%\include{chapters/ch_theory}
\include{chapters/ch_experiment}
\include{chapters/ch_data_analysis}
\include{chapters/ch_data_analysis_2}
\include{chapters/ch_results}
\include{chapters/ch_summary}

@@

\appendix
%\addappheadtotoc
\include{chapters/ch_appendix}

### 2020-07-16 chapters/ch_data_analysis.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

\chapter{Data Analysis I: Calibration and Corrections}

@@

\begin{figure}[!htb]
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
        \caption{Y-residual distribution for all pads before correction.} \label{fig:yoff_allBefore}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
        \caption{Y-residual distribution for all pads after correction.} \label{fig:yoff_allAfter}
    \end{subfigure}
    \caption{ }
\label{fig:yoff}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{mcresponse.png}
        \caption{200 MeV/c $\pi^-$ MC digitized track} \label{fig:mcevent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{event5.png} 
        \caption{Experimental data event.} \label{fig:dataevent}
    \end{subfigure}
    \caption{}
\label{fig:mcDataEmbedtrack}
\end{figure}

@@

    \caption{ }
\label{fig:132_eff}
\end{figure}

@@

    \caption{ }
\label{fig:108_eff}
\end{figure}

>>>>

\chapter{Data Analysis I: Calibration and Corrections}
\label{chap:data1}

@@

\begin{figure}[!htb]
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
        \caption{Before applying the timing correction.} \label{fig:yoff_allBefore}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
        \caption{After applying the timing correction.} \label{fig:yoff_allAfter}
    \end{subfigure}
    \caption{Y-residual distributions for all pads.}
\label{fig:yoff}
\end{figure}

@@

\begin{figure}[!htb]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{mcresponse.png}
        \caption{200 MeV/c $\pi^-$ MC digitized track} \label{fig:mcevent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\linewidth]{event5.png} 
        \caption{Experimental data event.} \label{fig:dataevent}
    \end{subfigure}
    \caption{Charge readout of the pad plane, showing the max ADC in a given pad, for both the MC simulation and experimental data.}
\label{fig:mcDataEmbedtrack}
\end{figure}

@@

    \caption{Pion efficiencies in the $\tin{132}{124}$ system. }
\label{fig:132_eff}
\end{figure}

@@

    \caption{ Pion efficiencies in the $\tin{108}{112}$ system. }
\label{fig:108_eff}
\end{figure}

### 2020-07-16 chapters/ch_data_analysis_2.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

\section{Beam Particle Identification}

@@

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterBottom.png} 
        \caption{Side view of the TPC showing the edge effect near the bottom.} \label{fig:clusterBottom}
    \end{subfigure}
\caption{}    
\label{fig:edge}
\end{figure}

@@

Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems. Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}. This helped to eliminate some of the poorly reconstructed beam events. The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionspectra}.

@@

\begin{figure}[!htb]
\centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn108.png} 
        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn112.png} 
        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn124.png} 
        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn132.png} 
        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
    \end{subfigure}
\caption{}
\label{fig:vertexdist}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
         \caption{$y=x$}
         \label{fig:pim108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:pim132angle}
     \end{subfigure}
        \caption{Three simple graphs}
        \label{fig:pim}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
         \caption{$\tin{108}{112}$ $\pi^-$ angular spectra.}
         \label{fig:pip108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:pip132angle}
     \end{subfigure}
        \caption{Three simple graphs}
        \label{fig:pip}
\end{figure}

>>>>

\section{Beam Particle Identification}
\label{sec:beam}

@@

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{clusterBottom.png} 
        \caption{Side view of the TPC showing the edge effect near the bottom.} \label{fig:clusterBottom}
    \end{subfigure}
\caption{Example showing the edge effect has on the cluster positions near the extreme edges of the TPC volume. }    
\label{fig:edge}
\end{figure}

@@

Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems. Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}. This helped to eliminate some of the poorly reconstructed beam events. The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionSpectra}.

@@

\begin{figure}[!htb]
\centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn108.png} 
        \caption{$\tin{108}{124}$ system.} \label{fig:vertex108}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn112.png} 
        \caption{$\tin{112}{124}$ system.} \label{fig:vertex112}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn124.png} 
        \caption{$\tin{124}{112}$ system.} \label{fig:vertex124}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{vertex-Sn132.png} 
        \caption{$\tin{132}{124}$ system.} \label{fig:vertex132}
    \end{subfigure}
\caption{Vertex distributions, around the target region, of all experimental runs for each secondary beam system. }
\label{fig:vertexdist}
\end{figure}

@@

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
         \caption{$\pi^-$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pim108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
         \caption{$\pi^-$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pim132angle}
     \end{subfigure}
        \caption{Measured $\pi^-$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
        \label{fig:pim}
\end{figure}

\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
         \caption{$\pi^+$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pip108angle}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
         \caption{$\pi^+$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
         \label{fig:pip132angle}
     \end{subfigure}
        \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
        \label{fig:pip}
\end{figure}

### 2020-07-16 chapters/ch_experiment.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

%explain in breif the basics of a time projection chamber 
A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors CITE HERE. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis. Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage. The field cage holds the detector gas and also defines a constant electric field that is typically aligned with the magnetic field of a magnet that surrounds the TPC. In this Spirit TPC, both these electric and magnetic fields are vertical.

@@

\begin{figure}[!htb]
\includegraphics[width=\textwidth]{fc_overview.png}
\label{fig:fc_overview}
\caption{Overview of the field cage mechanical components.}
\end{figure}

@@

Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem. The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated as,
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

\caption{}
\label{tb:gasprop}
\end{table}

@@

There were two beam drift chambers (BDC) located along the beam pipe upstream of the magnet and after the last focusing quadropole magnet. These beam drift chambers contained two sets of parallel plate avalanche counters (PPAC) which could get the (x,y) coordinates of the beam. Since they were places approximately \SI{1}{\metre} apart, we were able to track the direction of the beam in the beam line using a linear extrapolation. The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}. The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found. The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Sections~\ref{sec:beamangle,sec:pionSpectra}.

>>>>

%explain in breif the basics of a time projection chamber 
A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector. TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}. In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis. Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage. The field cage holds the detector gas and also defines a constant electric field that is typically aligned with the magnetic field of a magnet that surrounds the TPC. In this Spirit TPC, both these electric and magnetic fields are vertical.

@@

\begin{figure}[!htb]
\includegraphics[width=.9\textwidth]{fc_overview.png}
\label{fig:fc_overview}
\caption{Overview of the field cage mechanical components.}
\end{figure}

@@

Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem. The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated as,
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

\caption{Gas properties of P-10 gas at 1 atm pressure.}
\label{tb:gasprop}
\end{table}

@@

There were two beam drift chambers (BDC) located along the beam pipe upstream of the magnet and after the last focusing quadropole magnet. These beam drift chambers contained two sets of parallel plate avalanche counters (PPAC) which could get the (x,y) coordinates of the beam. Since they were places approximately \SI{1}{\metre} apart, we were able to track the direction of the beam in the beam line using a linear extrapolation. The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}. The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found. The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Section~\ref{sec:beamangle} and Section~\ref{sec:pionSpectra}.

### 2020-07-16 chapters/ch_intro.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

\chapter{Introduction}
On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations, reinvigorating the question about the nature of dense matter in the universe.

@@

Without a balancing force the Coulomb force between protons would render the nucleus unstable. This balancing force is called the strong nuclear force which exerts a large attractive force over a very short range. The strong force itself is the fundamental interaction between the fundamental quarks which make up the nucleons. At low energies the quark structure of nucleons is less important and nucleons can be thought of as fundamental particles. The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress. It is for this reason that the distance between nuclei is a near constant value, and therefore the density distribution over a wide range of nuclei is remarkably constant CITE HERE. This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.

Due to the nature of the strong force, nuclei can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility. This picture was remarkably successful at describing the binding energies of normal nuclea matter at saturation density. The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:

\begin{equation}
\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
\label{eq:semiEmp}
\end{equation}

@@

\section{Density Dependence of the Symmetry Energy}
In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ CITE HERE. In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$. Figure~\ref{fig:symDen} shows some of the experimental constraints which have been performed by a series of independent measurements and observables \cite{awayforward}. Typically the density dependence of the symmetry energy can be described through an effective interaction used to describe the phenomenological observations of nucleon-nucleon interactions observed in nuclei. One of such interactions is the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations \cite{skyrme}. Several Skyrme parameterizations are shown as lines in Fig.~\ref{fig:symDen}. Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars. The goal of the nuclear EoS community over the last decade have been particularly focused on constraining the symmetry energy at these high densities.

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=.5]{constraints}
\caption{Experimental constraints of the density dependence of the symmetry energy taken from \cite{awayforward}}
\label{fig:symDen}
\end{figure}

>>>>

\chapter{Introduction}
On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations \cite{ns_2017}, reinvigorating the question about the nature of dense matter in the universe.

@@

Without a balancing force the Coulomb force between protons would render the nucleus unstable. This balancing force is called the strong nuclear force which exerts a large attractive force over a very short range. The strong force itself is the fundamental interaction between the fundamental quarks which make up the nucleons. At low energies the quark structure of nucleons is less important and nucleons can be thought of as fundamental particles. The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress. It is for this reason that the distance between nuclei is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the heavist to the light \cite{krane}. This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.

Due to the nature of the strong force, nuclei can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility. This picture was remarkably successful at describing the binding energies of normal nuclea matter at saturation density. The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $E/A$:

\begin{equation}
\frac{E}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
\label{eq:semiEmp}
\end{equation}

@@

\section{Density Dependence of the Symmetry Energy}

\begin{figure}[!htb]
\centering
\includegraphics[scale=.3]{densityDep.png}
\caption{Current experimental constraints on the density dependence of the symmetry energy. The measured data points are discussed in the text. The 4 polygonal and shaded regions are constraints from: 1.) pion analysis from Ref. \cite{xia2009,xie2013}, 2.) a separate pion analysis from Ref. \cite{feng2010}, 3.) ASYEOS experiment \cite{russo2011, russo2016, cozma2013} and 4.) isobaric analog states \cite{dan2014} }
\label{fig:symDen}
\end{figure}

In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ \cite{neutronstar}. In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$. Figure~\ref{fig:symDen} shows the current status of laboratory constrains on on the density dependence of the symmetry energy. The red region labeled 4.) for $\rho < \rho_o$, corresponds to constraints from isobaric analog states \cite{dan2014}. Notice the symmetry energy is consistent across  a series of independent measurements and observables, but not at higher densities \cite{awayforward}. At higher densities, the two green sections correspond to constriants obtained from total pion yields, and pion ratio $\pi^-/\pi^+$, from Au + Au collisions at various beam energies \cite{fopi}. Region 1.) is obtained from analysis performed in References~\cite{xia2009,xie2013}. Region 2.) corresponds to Reference~\cite{feng2010}. The blue shaded region (labeled 3.) corresponds to constraints obtained through proton and neutron elliptical flow reported in References~\cite{russo2016,cozma2016,cozma2017}.

 Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars. The goal of the nuclear EoS community over the last decade have been particularly focused on constraining the symmetry energy at these high densities.

@@
